For each part of the calculation outlined above: describe the code that exists within GAMBIT and/or the ready-made interface to a standard HEP code. Show validation plots (with possible links to appendices).

\subsection{Cross-section calculations}

\begin{itemize}
\item Describe fast SUSY cross-section calculator design and implementation
\item Link to paper if we have an external paper
\item Include validation plots
\item Describe parameters that can be varied in yaml file.
\item Describe how we treat cross-sections for generic models.
\end{itemize}

\subsection{Monte Carlo event generation}

\begin{itemize}
\item Describe existing Pythia 8 interface (and OpenMP parallelisation)
\item Explain how to interface a different MC generator?
\item Describe parameters that can be varied in yaml file for Pythia.
\end{itemize}

\subsection{Detector simulation}

\begin{itemize}
\item Describe DELPHES interface for ATLAS and CMS processes
\item FastSim: Give brief description of simulation, with link to FastSim paper?)
\item Show some FastSim validation plots? (not really necessary if we have a FastSim paper, but might be useful to show DELPHES vs FastSim plots with speed comparisons)
\item Describe parameters that can be varied in yaml file.
\end{itemize}

\subsection{Event analysis framework}

\begin{itemize}
\item Describe event analysis framework
\item Emphasise applicability to simulated and/or reconstructed events
\item Describe the LHC analyses included in the module and include some validation plots showing some nicely reproduced ATLAS and CMS limits.
\item Explain how to add new analyses
\end{itemize}

\subsection{Statistics}

\begin{itemize}
\item Describe the various likelihoods one may calculate given a signal yield (overlaps with main GAMBIT framework- perhaps we need to release a small stats offshoot within ColliderBit for standlone users?))
\item Describe parameters that can be varied in yaml file.
\end{itemize}
